\documentclass[12pt]{article}

% House style preamble
\input{.house-style/preamble.tex}

% Print unpublished manuscript titles like article titles rather than standalone
% works.
\DeclareFieldFormat[unpublished]{title}{#1\isdot}

% Update PDF metadata
\hypersetup{
  pdftitle={Kinds as Projectibility Profiles: Support Grades and Demotion Rules},
  pdfkeywords={natural kinds, projectibility, homeostatic property clusters, kind claims, philosophy of science}
}

\title{Kinds as Projectibility Profiles:\\Support Grades and Demotion Rules}
\author{Brett Reynolds \orcidlink{0000-0003-0073-7195}%
\thanks{Contact: \href{mailto:brett.reynolds@humber.ca}{brett.reynolds@humber.ca}}\\
Humber Polytechnic \& University of Toronto}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Anti-essentialist theories explain how categories can be real without essences.
Homeostatic property cluster theory gives a Boydian answer, but
\term{homeostatic} now covers achievements that should be kept apart: stable
projectible profiles, network order, maintenance, and corrective control. This
paper gives anti-essentialist kind theory a stricter support-grade and demotion
vocabulary. A
category earns standing by tracking a projectibility profile: a worldly pattern
whose partial observation licenses inferences under specified conditions.
Profile analysis asks what the category projects, what supports the projection,
what orders and stabilizes it, and where the claim should be demoted or rejected.
Homeostatic control is a subcase, not the default.
\end{abstract}

\noindent\textbf{Keywords:} natural kinds; projectibility; homeostatic property clusters; kind claims; philosophy of science

\section{Introduction}

Anti-essentialist kind theory begins from a familiar pressure. Many categories
that matter to science, grammar, social inquiry, and ordinary reasoning don't
share an essence. Their members don't need to have one property in common, and
their internal organization may show gradience, historical contingency, or
context-sensitivity.
Still, some of these categories support reliable inference. They let us project
from observed features to further expectations: to predict, explain, intervene,
measure, compare, or learn.

Boyd's homeostatic property cluster framework offers one influential route
through this pressure.
A kind can be real when a cluster of properties supports induction and
explanation because category use has been accommodated to worldly structure
rather than because every member shares a defining essence
\citep{boyd1991,boyd1999}. That insight remains the starting point. But the word
\term{homeostatic} has been asked to do too much. In different contexts it can
name a stable profile, a network of relations, a maintaining process, or a
corrective-control system. The stricter use proposed here is a regimentation of
loose post-Boyd usage, not an exegetical claim that Boyd meant only corrective
control.

Those aren't the same achievement. A profile can be stable without a known
maintainer. It can be ordered without being maintained. It can be maintained
without being controlled. And it can support useful projection without earning
the stricter label \term{homeostatic}. Slater's stable property cluster account
presses the stability point \citep{slater2015}; Khalidi's causal-network account
presses the order point \citep{khalidi2013,khalidi2018nodes}; Weinberger's
account of homeostasis and causal control presses the control point
\citep{weinberger2026HomeostasisCausalControlBP}. The lesson isn't that Boyd
was wrong, but that the evidential work needs sharper labels.

Call the label \term{projectibility profile}. A profile, in this use, is the
worldly property-and-relation pattern that makes partial observation a guide to
further expectation within a stated range. The analysis of a profile specifies
that pattern: which projection it supports, how we diagnose it, what makes the
projected-from and projected-to properties co-available, what stabilizes that
co-availability, and where the projection fails.

I use \term{category} for the sorting item, \term{profile} for the worldly
pattern that supports projection, and \term{kind claim} for the assertion that
the category earns standing by tracking such a profile. I use
\term{kind} inclusively for any projectibility profile. In this minimal use,
\term{kind} isn't an honorific metaphysical status; it names a standing claim
that a category tracks a profile supporting a declared projection. A
\term{natural kind} claim and a \term{homeostatic} claim are stronger, separable
assertions about that support. The inclusive use doesn't make kindhood cheap. It
shifts the burden from a single kind/non-kind threshold to what the category
projects, under what scope, with what support, and with what demotion
conditions.

Projectibility comes first methodologically and explanatorily. We start by
asking what the category lets us infer, then ask what profile supports that
inference, what relations make the profile non-accidental, what stabilizes it,
and where it fails. That order doesn't imply that projectibility is
metaphysically prior to every stabilizing structure. A mechanism earns its place
in a kind claim by explaining why the declared projection holds, not by being
named.\footnote{This paper doesn't take a stand on the stronger constitutive
thesis sometimes associated with \term{projectibility-first} formulations.
Nothing in the argument depends on that stronger claim.}

The contribution is primarily regulative. This paper doesn't add a new entity to
the ontology of kinds; it gives existing anti-essentialist resources a
support-grade and demotion discipline. The order of use is: identify the
projection, specify the profile, identify the ordering relation, state what
stabilizes the profile, and declare scope and demotion conditions. It also keeps
explanatory levels apart: a corpus distribution, a social practice, and a
cognitive mechanism can't trade explanatory roles without an argued bridge.

Closest to this account is Ereshefsky and Reydon's grounded functionality
account \citep{ereshefsky2023gfa}. It ties a kind's standing to the aims it's
posited for, and it blocks conventionalism with a world-side condition: a
proposed category counts only when the relation it assumes is grounded in some
aspect of the world, not merely in our conceptions of it. It also lets the
function fix which aspect of the world must answer to it. On those structural
questions, the accounts largely agree.

Where they part company is the licensed function. The disagreement isn't over
whether functions must be grounded. It's over which grounded functions earn kind
standing. Inference here is broader than induction over shared
properties.\footnote{\textcite{ereshefsky2023gfa} use stable microbial
re-identification and descent-based taxonomy against an induction requirement.}

Stable re-identification and historical individuation can still be projections:
a marker can point to an identified group, and membership can point to shared
ancestry. With the function read that wide, fixing it as inference doesn't
exclude Ereshefsky and Reydon's non-inductive cases; it redescribes them. The
real contrast, then, is between sorting and projection: a class sorts, while a
kind in good standing licenses projection. A grounded, aim-meeting category that
licenses no further expectation is at most a thin class.

Two further gains follow. Their single grounding condition is split into an
order axis and a stabilization axis: stabilizer, maintainer, and controller. The
natural/non-natural boundary is replaced by support tiers and declared scope. A
conventionally stabilized category such as \term{Canadian permanent resident} is
a scoped kind here, projectible within its institution and demoted by scope
overreach outside it. In that respect, the proposal accepts the
multidimensional diagnosis in \textcite{ludwig2018LettingGo,ludwig2024WithoutNaturalKinds}
but rejects the proposed cure: instead of letting go of kind talk, disaggregate
the supports that kind talk has been used to claim.

A statistical cousin of this scope discipline appears in the literature on
external validity. \textcite{egami2023elements} decompose generalization into
populations, treatments, outcomes, and contexts, each relative to a target the
inquiry fixes, so that stating a target is stating a scope condition. The profile
view shares that target-relativity but adds an axis they hold fixed: their
framework asks whether a fixed category's effect carries to a new setting, while
profile analysis also asks whether the projection survives when we change what
counts as the category.

\section{The Profile Package}

A profile specification answers six questions. First, what's the projection
target? Second, what pattern of properties and relations supports it? Third,
what evidence diagnoses that pattern? Fourth, what ordering relation makes
source and target properties co-available rather than accidental
co-occurrences? Fifth, what stabilizes that co-availability across relevant
contexts, timescales, or perturbations? Sixth, what scope conditions and
failure modes mark the projection's limits?

\subsection{Projection and Profile}

Recurrence is cheap. A list of co-occurring properties becomes theoretically
interesting only when it licenses some further expectation. The expectation may
be predictive, explanatory, taxonomic, or intervention-guiding. It may concern a
biological process, a grammatical pattern, a measurement practice, or a social
role. But without a declared projection target, the category remains a
description of observed association.

By \term{profile}, I mean the property-and-relation pattern whose partial
observation licenses the inference. The licensing runs in a direction: from
observed features to some further feature, contrast, intervention,
re-identification, or decision. A symmetric association that supports no directed
expectation isn't yet a projection in this sense. Merely sorting familiar cases,
coordinating attention, or recording an association isn't enough.

The expectation also has to put the claim at risk. If the advertised target
only restates the sorting rule, or follows by stipulation from the way the label
was assigned, the category may be useful without yet supporting a kind claim in
good standing. Projection requires some further expectation that could fail.

The direction and scope also have to be specifiable independently of the
successful cases. A projected target, contrast, and scope can't be drawn around
exactly the observations that already fit. Otherwise the proposal is a fitted
grouping, not a profile with projectible standing.
This condition guards against gerrymandering; it isn't an answer to Goodman's
problem by itself. A grue-like predicate can be specified in advance. The
anti-Goodman work belongs to the ordering requirement, which asks whether the
projected relation is a genuine dependence rather than a case-fitting rule.

The same pattern appears at different scales. In biology, observing juvenile form
may license expectations about adult morphology, development, ecology, or
reproduction. In grammar, seeing the verb form \mention{written} licenses
expectations about \mention{has written}, while seeing \mention{drew} licenses a
different surrounding frame. In a social case, a permanent-resident status marker
licenses expectations about entitlements, sanctions, and repair paths. The
pattern doesn't need to be one-level or one-mechanism.

Diagnostics come next, and they belong to the specification, not to the profile
itself. They're the tests, traces, contrasts, measurements, or judgments by
which inquiry locates the profile and checks whether a proposed projection
travels beyond the cases that introduced it.

\subsection{Order and Stabilization}

Non-accidental co-availability poses the structural question. Khalidi's causal
network account supplies one important model: kinds aren't mere concatenations
of properties, but nodes in ordered networks where some properties help explain
others \citep{khalidi2018nodes}. In other domains, the relevant order may be
functional, conventional, normative, or algorithmic rather than narrowly causal.
Pluralism enters the explanatory description, not the evidential demand.
Functional, conventional, and normative relations still have to be worldly
dependences; the terms name forms that dependence can take, not weaker grades of
support.

Here the issue is synchronic or structural: why are the
projected-from and projected-to properties co-available in the cases at issue?
In the broader use, the ordering relation still has to explain the declared
projection; mere cue status isn't enough. Where the stronger Khalidian
natural-kind claim is at stake, conventional or normative order needs causal,
causal-historical, or causal-normative support.

Different forms of order answer this demand differently. In causal cases, the
target property depends on processes that also explain why the source features
are reliable guides. In conventional or institutional cases, the dependence can
be constitutive: \term{Canadian permanent resident} status is partly constituted
by records, rules, and recognition practices that make entitlements, sanctions,
and repair procedures follow. Those practices don't merely cue the status; they
help make the status and its consequences the case. A cue remains diagnostic
when it tracks the target through a background condition without entering the
dependence that explains the projection.

Functional and algorithmic cases need the same discipline. A component belongs
to the ordering relation when it enters the organization that produces the
projected target. Retrieval in a large\mbox{-}language\mbox{-}model pipeline explains citation
accuracy only if live records enter the production route that yields the citation
or refusal. If retrieval merely correlates with easier tasks or cleaner
benchmarks, it remains a diagnostic cue.

The positive test is counterfactual within the declared level. Hold the cue
fixed and vary the alleged ordering relation: if the projection changes in the
predicted direction, the relation has earned support; if it doesn't, the cue was
only diagnostic.

In institutional cases, a record without recognition,
enforcement, or repair may document a status without making its entitlements
follow. In algorithmic cases, stale retrieval or retrieval detached from
generation may mark an easier task without ordering a truth-tracking route.

Where direct intervention on the ordering relation isn't available, as in many
historical or grammatical cases, the test is indirect: held-out constructions,
dialectal or acquisition contrasts, and predicted dissociations have to change as
the proposed order says they should. The claim is weaker if the evidence tests
only the profile's reach, not the ordering relation.

The level has to be marked. Otherwise a social coordination pattern, a cognitive
mechanism, and a corpus distribution begin to explain one another by
equivocation.

The next question is \term{stabilization}: what keeps the profile available
across the relevant range of contexts, timescales, and perturbations? Some
answers to the structural question also answer the stabilization question, but
the claims differ. Network order says why co-availability isn't coincidence;
stabilization says why it persists or recurs within the scope of the projection.
Sometimes no separable
stabilizer has been identified. Robust stability can still support projection
across its demonstrated range, but claims about maintenance or control require
additional evidence.

Identifying a stabilizer adds modal information, not prestige. A stable profile
is a complete success at its grade when held-out cases within the declared range
continue to support the projection. A stabilizer claim is earned only by
contrasts in which varying the proposed stabilizer preserves, weakens, or breaks
the projection in predictable ways.

A \term{stabilizer} explains why co-availability persists or recurs across the
declared range. It may be a relation, constraint, mechanism, practice, or shared
standard. Stabilizers don't need to be active processes: selectional constraints,
institutional practices, and ecological regularities can all stabilize a
profile. What matters is that they explain why observing part of the profile
gives reason to expect the rest.

A \term{maintainer} is more specific. It's an active stabilizer that preserves
or regenerates the profile over time or under perturbation. A maintainer may be
developmental, social, institutional, biological, or interactional. It can keep a
profile available without registering each departure as an error or returning a
system to a fixed value.

A \term{controller} is more specific again. It's a maintainer with
perturbation-sensitive organization that preserves a higher-scale relation
through lower-scale variation. On the stricter use defended in
\textcite{reynolds2026notEveryStableCluster}, the term
\term{homeostatic} belongs here. Homeostasis isn't a synonym for recurrence,
order, or maintenance. It marks the demanding case where corrective organization
preserves or restores a relation despite perturbation. The restriction fits
Weinberger's recent account of homeostasis and causal control, where feedback
systems use lower-scale loops to produce higher-scale causal relations rather
than simply holding variables constant
\citep{weinberger2026HomeostasisCausalControlBP}.

Readers who reserve \term{homeostatic} for Boyd's broader maintained clusters
can read the strict label as the controller grade throughout. The substantive
claim is that controller-grade evidence shouldn't be presumed from stability,
network order, or maintenance alone.

One contrast keeps the levels apart. A valley's shape may be stabilized by
erosion and gravity; wound closure is maintained by active healing; body
temperature is controlled by corrective feedback. Only the last case is
homeostatic in the strict sense. The schema is simple: order explains
non-accidental co-availability; stabilization explains persistence or recurrence;
maintenance is active preservation or regeneration; control is
perturbation-sensitive maintenance of a higher-scale relation through
lower-scale variation.

\subsection{Scope and Failure}

Scope and failure complete the specification. A profile is projectible only
across a declared range of conditions and contrasts, for a declared inferential
purpose. The claim is field-relative in this limited sense: the inquiry selects
which available projections matter, but it doesn't make them available. When the
profile supports one family of inferences but not another, the result is a scoped
claim: localize the projection, mark its defeats, and don't export it as a
broader kind claim.

\section{Why Projectibility Comes First}

Goodman's new riddle made \term{projectibility} the name for a hard problem:
some predicates that fit observed cases still don't license projection to
unobserved ones \citep{Goodman1955}. This use of \term{projectible} is Boydian
rather than Goodmanian in one central respect. It doesn't treat projectibility
primarily as entrenchment in our predicate-using history. It asks what worldly
order makes a category a reliable guide to further expectation.

This doesn't dissolve Goodman's problem. It presupposes a field-specific
distinction between grounding order and case-fitting pattern, and it inherits
local standards of explanation, dependence, and defeat. The aim is to regiment
candidates, not to certify projectibility from outside the special sciences.

The regimentation has two parts. The independence rule is prospective: a
candidate profile has to be specified before the successes are counted and then
carried to held-out cases or perturbations. That rule blocks scope
gerrymandering; it doesn't by itself rule out a grue-like predicate. The ordering
requirement carries the anti-Goodman burden: the projection has to answer to a
field-licensed dependence rather than a case-fitting rule.

A benchmark regularity can be genuine without being projectible: the observed
successes may cluster because the test cases share an accidental background
feature rather than because the advertised relation supports the projection.
That regularity doesn't support the kind claim unless it projects to fresh cases
where the same support relation is still operative.

That realist turn is the point of Boyd's anti-essentialism. We don't need a
shared essence to explain why a category supports induction. We need a
non-accidental profile that our category use has partly learned to track
\citep{boyd1991}.

What remains Boydian in the stricter vocabulary is this accommodation claim:
category use answers to worldly structure that makes inference reliable. The
stricter vocabulary narrows \term{homeostatic}, not that realist core. A stable or
network-ordered profile can be fully Boydian without being a controller.

Projectibility is the target of the kind claim. Mechanisms, networks,
histories, standards, and practices matter because they explain why projection
works.
This order blocks a common slide. Recurrence gets promoted to stability,
stability to stabilization, stabilization to homeostasis, and homeostasis to
kindhood. The profile vocabulary reverses that order. Start with the inference.
Ask what profile licenses it. Then ask which stabilizing claims have actually
been earned.

Starting with a mechanism makes the slide easier. Once a process has been named,
the temptation is to let it certify the kind before asking what projection it
improves. The profile-first order forces the mechanism to answer a narrower
question: what inference would become weaker if that mechanism were removed?

Mechanisms may be primary in the world. During inquiry, though, we may have good
evidence that a profile is projectible before we know what maintains it. The
reliable projection still depends on whatever worldly relations make the profile
non-accidental. Epistemic order and grounding order can come apart. The proposed
discipline is to avoid treating the unknown ground as if it had already been
identified, named, and promoted to homeostasis.

Stable property clusters fit the account for the same reason. Slater is right
that a natural-kind claim doesn't need to wait for a named homeostatic mechanism,
and his stability demand is already projectibility-shaped: observing enough of a
subcluster should license expectations about the larger cluster
\citep{slater2015}. Stability can be enough for a scoped kind claim when it
licenses the declared projection. Scoped doesn't mean second-class: a stable
profile can be a complete success across its demonstrated range. The profile
question then asks which further support claims about stabilizers, maintainers,
or controllers have also been earned. Maintenance or control adds support-grade
information, and may extend what can be projected; it doesn't make a stable
success more real. A category may support modest local projections while failing
to support broader ones. Both results are informative.

Purpose-relativity enters here with care. A profile is projectible for a family
of questions, contrasts, and evidential practices. In physiology, a useful
profile may license intervention and measurement. In grammar, it may license
expectations about distributions, manipulations, and borderline cases. In social
ontology, it may license expectations about uptake, status, sanctions, or
institutional repair. The practice selects which projection matters. It doesn't
make the projection true.

Availability is relative to operative stabilizers, which may be causal,
conventional, or normative. Strong mind-independence isn't required: an
institutionally stabilized profile can be available even when its stabilizers
depend on us. What the account refuses is availability created by the inquirer's
interest alone. \term{Worldly} doesn't mean human-independent. It means that,
given the relevant practices, standards, mechanisms, or institutions, the
projected relation isn't made available merely by the inquirer's decision to
group the cases together.

A projectibility profile differs from a checklist. Checklists record
features. Profiles connect observed features to directed expectations under
scope conditions. A checklist asks whether an item has properties \(P_1\),
\(P_2\), and \(P_3\). A profile asks whether observing \(P_1\) and \(P_2\), in
the relevant setting, licenses expectation of \(P_3\) rather than merely
recording that all three co-occur. It also asks what would defeat that
expectation.

That difference matters for explanatory economy. A mechanism-first paper can
finish by naming structures that hold a cluster together. A profile-first paper
has to say what those structures let us project, what failures would demote the
claim, and whether the advertised stabilizer operates at the right explanatory
level. The contribution is a discipline for using categories as guides to
further expectation.

Here's a simple test for mechanism talk. If removing the mechanism leaves the
same predictions in place, the mechanism hasn't earned its role in the kind
claim. It may still be a true description of the system, but it's not yet doing
the explanatory job a projectibility-profile analysis needs.

\section{Profiles Across Levels}

Profile vocabulary should travel without making every domain look biological.
Profiles can be distributional, cross-linguistic, epistemic, behavioural, or
social-practical, among others. This section applies the vocabulary at different
levels. Section~5 returns to what goes wrong when evidence at one level is
projected to another without a bridge claim. The cases function as portability
tests: they show what the audit asks, while the companion papers defend the
underlying claims.

My work on English definiteness supplies a concrete linguistic case. There,
\term{deitality} names a morphosyntactic profile identified by distributional
tests involving noun phrases introduced by the definite article \mention{the},
demonstratives, possessives, and some proper-name uses. The point isn't to
rename every semantically definite expression. It's to ask whether those forms
support further expectations: which diagnostics should cluster, where form and
referent identifiability should come apart, which contrasts should shift under
prosody or specificity manipulation, and which variation should appear across
dialects and acquisition \citep{reynolds2025definitenessProjectibility}. The
category earns its place by those projections.

The audit doesn't require the full diagnostic list; one familiar contrast shows
the relevant kind of projection. Forms such as \mention{drew} and
\mention{written} both concern past events, while they project different
grammatical surroundings
(\mention{I drew it} and \mention{I have written it}, rather than
\mention{*I have drew it} or \mention{*I written it}). The deitality claim is
similar. Seeing \mention{the}, \mention{this}, \mention{my}, or some proper-name
uses should guide expectations about where the expression can appear, when
prosody or specificity will change the judgment, and where weak readings such as
\mention{go to the hospital}, dialectal variation, or acquisition should split.
The acquisition claim is a derived conjecture rather than an established result.

That is enough to show how the audit runs. The ordering relation is
grammatical-distributional. The support grade, if earned, is stabilization
through grammaticalization, discourse pressure, and iterated transmission, not
corrective control. The scope is English uses of \mention{the}, demonstratives,
possessives, and related noun-phrase patterns; proper names, weak-definite
frames, dialects, and cross-linguistic extensions are separate tests, not
automatic exports. The claim is demoted if the diagnostics only sort familiar
forms, or if the predicted dissociations collapse into semantic definiteness
alone.

Cross-linguistic comparison follows the same discipline. Suppose a database uses
one label to compare similar constructions in several languages. The label may
remain an instrument for measurement. It licenses a stronger kind claim only if
knowing the label helps us expect something beyond the table already coded: how a
new language patterns, which neighbouring meanings pattern together, or where the
comparison should break down \citep{reynolds2025comparanda}.

For example, if several constructions are coded as \term{passive}, the stronger
claim is not just that coders applied the label consistently. It's that the
label helps predict further facts: which participant becomes subject-like,
whether the agent can be omitted, which morphology clusters with the
construction, and where the comparison stops travelling.

Here the stabilizing background may include processing, diachrony, discourse
pressure, and literate or institutional transmission. Acquisition can also
stabilize a cross-linguistic profile, but only if acquisition pressures help
explain why the profile recurs across languages. Measurement design supplies the
audit trail, not a stabilizer. None of those should be collapsed into a universal
language-internal category.

In linguistic typology, the comparison of grammatical patterns across languages,
travel means more than clean coding. Once the coding decision and sample are
fixed, the profile should predict held-out languages, constructions, or nearby
meanings that often pattern together. Failure may localize the claim by area,
genealogy, construction type, or coding purpose. If the apparent profile is
created by the coding manual or sampling grid, the result is a profile of
measurement practice, not a language-internal kind claim.

Large\mbox{-}language\mbox{-}model systems supply a different portability test. The candidate
kind claim concerns a \term{truth\mbox{-}tracking production route}: a
model-plus-task pipeline whose observed outputs license further expectations
about truth-relevant performance. Outputs are diagnostics here, not the bearer
of the kind claim.

Self\mbox{-}consistency, retrieval, and tool use are candidate supports for
particular projections. A profile analysis asks which projection each support
licenses: citation accuracy, correction after retrieval, calibrated refusal, or
robust transfer under task variation. The supports for a
truth\mbox{-}tracking claim are heterogeneous: output coherence, record access,
tool-mediated calculation or measurement, and calibration feedback
\citep{reynolds2026truthTrackingProfiles}. Because those supports can come
apart, a surface gain can leave open which kind claim the outputs support. The
kind claim, not the gain, is what licenses projection to new tasks.

The contrast is easiest to see with retrieval and self\mbox{-}consistency.
Retrieval lets the route consult records during production; self\mbox{-}consistency
compares the model's own sampled answers. Those interventions can improve
different projections: one may strengthen record use, while the other may stabilize
agreement among fluent outputs.

In document-grounding, the contrast is diagnostic. Retrieval and
self\mbox{-}consistency play two roles: they're candidate supports, and their
effects under test help locate the level of the candidate kind claim. If
retrieval changes citation accuracy because live records enter the production
route, the result is localization: a truth\mbox{-}tracking claim about a
record-backed route under that task profile, not about such systems in general.

If self\mbox{-}consistency raises agreement while
unsupported claims remain, the result is demotion: the route is
stabilized by coherence at the behavioural level, not truth\mbox{-}tracking.
Misclassifying that route as truth\mbox{-}tracking would license the wrong
projection to record-backed, tool-using, or out-of-distribution settings.
The mistake is a kind-claim mistake, not just a lower score: it predicts
transfer where only coherence has been stabilized.

A schematic contrast set shows what the audit would require. Fresh retrieval
should improve citation accuracy where stale retrieval doesn't; absent records
should license calibrated refusal rather than fluent invention; adversarial
documents should reveal whether the route follows records or prior familiarity;
and tool-enabled calculation should change failures that pure self-consistency
leaves in place. A route that improves agreement without those counterfactual
shifts is coherence-stabilized, not truth-tracking.

An evaluation profile should pre-register the level and support grade being
tested, the projection expected at that level, and the failure condition. The
result says whether the kind claim can stand as a truth\mbox{-}tracking
production route, should be localized to record-backed production, or should be
demoted to coherence\mbox{-}stabilized production in that setting.

In each case, the same vocabulary applies without imposing the same mechanism.
The \term{deitality} profile is grammatical and distributional. The
cross-linguistic profile is comparative and methodological. The
truth\mbox{-}tracking production-route profile is epistemic and process-level.
Calling all three \term{profiles} doesn't erase those differences. It lets us ask
the same disciplined questions: what gets projected, what pattern supports it,
what orders and stabilizes it, and where it fails.

Fields, practices, and institutions should be handled in the same
way.\footnote{\textcite[148]{boyd1999} already relativizes naturalness to a
\enquote{disciplinary matrix}, a family of inferential practices rather than an
ordinary academic domain. This account keeps that relativity but separates
Boyd's matrix from the social field or institution that carries the project.
Boyd also says that accommodation demands are \enquote{themselves homeostatically
related} \citep[157]{boyd1999}; here that support claim has to be earned by
evidence of corrective organization, not recurrence alone.}

They can be scope conditions or stabilizing backgrounds; in some social cases,
institutional facts may also belong to the profile itself. Fields and practices
shouldn't be promoted to kinds merely because they organize inquiry or action.
A field may condition which projections matter without itself being the
projection target. A practice may stabilize the profile associated with a
category without being the kind. Those are bridge claims, and they need their own
evidence.

\section{Failure Modes and Demotion}

Vocabulary is useful only if it can say no. Without failure modes,
\term{projectibility profile} would become another loose label for any
interesting cluster. A profile view needs demotion rules: ways to say that a
kind claim is too thin, projected at the wrong level, ordered by the wrong
relation, carrying unused mechanism talk, overpromoted, or too broad in scope.
The list is diagnostic rather than exhaustive. The modes track pressure points in
the six-question specification, but they can overlap when one error overgrades
support in more than one way.

The first failure mode is a thin profile. A thin profile has diagnostics that
identify familiar cases but don't license further expectations. In grammatical
work, the label \term{negative polarity item} can define a useful class:
\mention{ever} in \mention{I didn't ever go} and \mention{lift a finger} in
\mention{She didn't lift a finger} both need a licensing environment. The kind
claim is thin, though, if the label doesn't predict the distinct licensing
patterns of its members \citep{hoeksema2012}.

The second failure mode is wrong-level projection: evidence at one level is used
to project at another without a bridge. A corpus distribution can show that a
pattern is learnable from text, but not that a model stores, computes, or uses
that pattern internally. A social-practical regularity can show that a practice
coordinates action, but not a biological disposition in the people who
participate. A grammatical category in one language can motivate a comparison,
but not a profile that travels across languages. The projection may be possible;
the bridge has to be argued.

The third failure mode is misattributed order. This happens when a real
dependence is treated as underwriting the wrong inference. A stable or maintained
cluster may include a cue that tracks the target because both depend on a
background condition. A retrieved-document flag, for example, may track accurate
answers because both occur on easier tasks. The cue can help diagnose the
profile, but it doesn't explain the projected-to property. Here the order claim,
not the profile, is overdrawn.

The fourth failure mode is decorative mechanism. A paper names a mechanism, but
the mechanism doesn't underwrite any new projection. The test from Section~3
returns here as a demotion rule: if removing the mechanism talk leaves the same
predictions, the mechanism hasn't earned its place in the kind claim.

The fifth failure mode is over-homeostatization. Unlike decorative mechanism, the
support may contribute to projection. The error is grading it too high:
stability, network order, and maintenance shouldn't be relabelled as homeostasis
unless the evidence shows corrective organization preserving a higher-scale
relation through lower-scale variation. The same diagnostic applies when a
stabilizer is promoted to maintainer without evidence, but the homeostatic
version claims the strictest grade. A maintained convention, a stable
distribution, or a reliable measurement practice may be projectible without
being homeostatic.

The sixth failure mode is scope overreach. A profile can be real and useful
under one set of conditions while failing elsewhere. The error is to export the
projection without fresh tests. A category projectible for discourse analysis
may not be projectible for syntax. A model behaviour projectible under benchmark
conditions may not transfer to tool-using settings. A social status projectible
inside one institution may not travel across institutions. In cross-linguistic
comparison, a label that projects within a coding sample but can't generalize
beyond it has overrun its scope.

Demotion belongs inside realism. If kind claims are
answerable to the projections they support, then failed projection should change
the claim. Sometimes the right response is rejection: the category doesn't
support the advertised inference. Sometimes localization is right: the category
is projectible only under narrower scope conditions. Sometimes relevelling is
right: the pattern is real, but at a different explanatory level. Sometimes
support-grade demotion is right: the projection may stand, while the advertised
support is downgraded or removed. Misattributed order takes a parallel response:
keep the cue as diagnostic if it works, but demote the claim that it explains
the projected property. A controller becomes a maintainer, a maintainer becomes a
stabilizer, or a named mechanism drops out of the kind claim.

As with the admission condition above, localization earns its keep only when the
narrower scope is specifiable independently of the surviving cases: by domain,
register, condition, or contrast fixed before the failures were seen. A scope
drawn around exactly the successes redescribes failure as success; it doesn't
demote the claim.

Profile talk joins promotion and demotion. A category earns a better specified
claim when its profile supports reliable projection, when any claimed
stabilizers are identified at the right level, and when its limits are known. It
loses status when those conditions fail. The framework gives anti-essentialist
kind theory an exit procedure, not just a route into realism.

\section{Conclusion}

A kind claim earns standing when a category tracks a profile that
licenses projection under specified conditions, with the ordering and
stabilizing support the projection requires. The compact order is projection,
profile, support, scope. Each element matters. No single essence
need define the profile. Stabilization matters because recurrence needs
explanation. Projectibility matters because kind claims earn their scientific or
grammatical value by licensing further expectation.

Homeostasis remains important, but it no longer has to carry every achievement.
Two axes have to stay apart. Network order is a structural claim about why
co-availability isn't coincidence. Stabilization and maintenance concern
persistence and recurrence; control adds corrective return after perturbation. A
profile can be strong on one axis and weak on the other. Only the controller
case deserves the strict homeostatic label. Keeping those distinctions apart
preserves Boyd's anti-essentialist realism while incorporating Slater's
stability-first lesson and Khalidi's network discipline.

Several grounding questions remain open. The framework doesn't decide in
advance whether a projection is grounded causally, conventionally, normatively,
or by some mixture of supports. It doesn't solve Goodman's new riddle. And it
doesn't assume that all six profile questions will receive strong answers in
every case. It gives the order in which those answers have to be earned.

For recall, the vocabulary leaves five questions, with order and stabilization
named together rather than analytically merged. What does this category let us
infer? What pattern supports that inference? What orders and stabilizes the
pattern? What would defeat the projection? And when should the category be
demoted, localized, or rejected?

\section*{Acknowledgements}

I thank Ryan Briggs for discussion linking the scope conditions developed here to
the statistics of external validity and transportability. OpenAI Codex and Claude
Opus helped with drafting and editing. I'm responsible for all theoretical claims,
arguments, errors, and interpretive choices.

\clearpage
\begingroup
\setlength{\emergencystretch}{2em}
\printbibliography
\endgroup

\end{document}
